
---
title: "Example Vectra image analysis"
output:
  html_document:
    toc: yes
---


```{r load_package}
library(cowplot)
library(ggplot2)
source('spatstat_vectra.R')

plot_function <- function(csd, ctypes) {
    plt <- ggplot() + theme_bw() + 
           geom_point(data = csd[grep('Other', csd$Phenotype),],
               aes(x=x, y=y), col = '#b5b3ae', size=2.5) +
        geom_point(data = csd[csd$Phenotype%in%c('PAX5+PDL1-','PAX5+PDL1+'),],
               aes(x=x, y=y), col = '#fc8105', size=2.5) +
           geom_point(data = csd[grep('PAX5+PDL1+', csd$Phenotype, fixed=T),],
               aes(x=x, y=y), col = '#fff761', size=0.01, shape=1, stroke=1) + 
           geom_point(data = csd[grep('CD3+CD8-',csd$Phenotype, fixed=T),],
               aes(x=x, y=y), col = '#34c9eb', size=2.5) +
           geom_point(data = csd[grep('CD3+CD8+',csd$Phenotype, fixed=T),],
               aes(x=x, y=y), col = '#4634eb', size=2.5) +
            geom_point(data = csd[grep('PD1+', csd$Phenotype, fixed=T),],
               aes(x=x, y=y), col = '#fff761', size=0.01,shape=1, stroke=1) + 
           geom_point(data = csd[grep('CD163', csd$Phenotype),],
               aes(x=x, y=y), col = '#a834eb', size=2.5) +
           geom_point(data = csd[grep('CD163+PDL1+', csd$Phenotype, fixed=T),],
               aes(x=x, y=y), col = '#fff761', size=0.01, shape=1, stroke=1) 
    return(plt +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                   panel.background = element_blank()))
}

```


```{r example_images, fig.width=12, fig.height=6}

images = c(
    'exported_mIF/HO105-003_[11564,35679]_cell_seg_data.txt'
#    'exported_mIF/HO105-003_[11564,35679]_cell_seg_data.txt',
#    'exported_mIF/HO105-180_[15869,40356]_cell_seg_data.txt'
#    'exported_mIF/HO105-126_[15975,49075]_cell_seg_data.txt',
#    'exported_mIF/HO105-025_[9621,46812]_cell_seg_data.txt',
#    'exported_mIF/HO105-036_[10536,42112]_cell_seg_data.txt',
#    'exported_mIF/HO105-182_[8076,47954]_cell_seg_data.txt',
#    'exported_mIF/HO105-010_[15489,46493]_cell_seg_data.txt',
#    'exported_mIF/HO105-020_[9758,38701]_cell_seg_data.txt',
#    'exported_mIF/HO105-080_[9000,38682]_cell_seg_data.txt',
#    'exported_mIF/HO105-010_[14712,46842]_cell_seg_data.txt',
#    'exported_mIF/HO105-020_[10157,38113]_cell_seg_data.txt'
)

ctypepairs = list(
    c('CD3+CD8-PD1-', 'CD8+CD8-PD1+', 'CD3+CD8+PD1-', 'CD3+CD8+PD1+'),
    c('CD163+PDL1-', 'CD163+PDL1+', 'CD3+CD8+PD1-', 'CD3+CD8+PD1+'),
    c('CD163+PDL1-', 'CD163+PDL1+', 'CD3+CD8+PD1-'),
    c('CD3+CD8-PD1-', 'CD3+CD8-PD1+', 'PAX5+PDL1-'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('PAX5+PDL1+', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+'),
    c('CD163+PDL1-', 'CD163+PDL1+')
                  )


XposCol = 'Cell X Position'
YposCol = 'Cell Y Position'
PhenoCol = 'Phenotype'

#i=1
for (i in 1:length(images)) {
    plist = list()
    Table <- purrr::map_df(images[i], read_cell_seg_data, 
                       pixels_per_micron = "auto", remove_units=FALSE)

    # replace empty phenotype with "Other"
    Table$Phenotype[Table$Phenotype == ""] = "Other"
  
    # define csd for ppp
    csd <- as.data.frame(Table[, c(PhenoCol, XposCol, YposCol)])
    colnames(csd) = c('Phenotype', 'x',  'y')

    plt1 <- local({

        plt1 = plot_function(csd, c('Tumor', 'Tcell', 'Macrophage', 'Other'))
        #plt <- plt + ggtitle(images[i]) 
        plt1 <- plt1 + #ggtitle(images[i]) + 
          theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                axis.title.x=element_blank(), axis.text.x=element_blank(),
                axis.ticks.x=element_blank(), axis.title.y=element_blank(), 
                axis.text.y=element_blank(), axis.ticks.y=element_blank(),
                panel.background = element_blank()) + ggtitle('all cell types')# + scale_y_reverse()
        return(plt1)
    })

    #print(plt1)    
    xlims <- layer_scales(plt1)$x$get_limits()
    ylims <- layer_scales(plt1)$y$get_limits()

    csd2 <- csd
    plt2 <- plot_function(csd2[csd2$Phenotype%in%ctypepairs[[i]],], 
                         c('Tumor', 'Tcell', 'Macrophage', 'Other')) + 
           xlim(xlims) + ylim(ylims)# + scale_y_reverse()
    plt2 <- plt2 + 
          theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                axis.title.x=element_blank(), axis.text.x=element_blank(),
                axis.ticks.x=element_blank(), axis.title.y=element_blank(), 
                axis.text.y=element_blank(), axis.ticks.y=element_blank(),
                panel.background = element_blank()) + 
           ggtitle(paste0('focus: ', paste0(ctypepairs[[i]], collapse='/')))# + scale_y_reverse() 

    plot_row <- plot_grid(plt1, plt2)

    title <- ggdraw() +
        draw_label(
                   images[i],
                   fontface = 'bold',
                   x = 0,
                   hjust = 0
                   ) +
           theme(
                 plot.margin = margin(0,0,0,7)
                 )
    
    print(plot_grid(title, plot_row, ncol=1, rel_heights = c(0.1,1)))
}


```


```{r, eval=T, cache=T, warning=F, message=F}

phenotype_complete = c("CD163+PDL1-","CD163+PDL1+",
                       "CD3+CD8-PD1-", "CD3+CD8-PD1+", "CD3+CD8+PD1-", "CD3+CD8+PD1+",
                       "Other","Other PDL1+",
                       "PAX5+PDL1-", "PAX5+PDL1+",
                       "")

color_complete = list("magenta", "brown",
                   "red", "blue", "green", "yellow",
                   "gray", "pink",
                   "orange", "cyan",
                   "gray")
names(color_complete) = phenotype_complete

fig.prefix.complete = './complete'
selected_spatstat_statistics = c('K', 'L') 
# can be any subset of ('K', 'Kdot', 'L', 'Ldot', 'F', 'J', 'G', 'Gdot', 'pcf')
r_vec = c(10)

plotter_sample = T
plotter_quadrat = T
plotter_alltypes = T

plotter = list(plotter_sample, plotter_quadrat, plotter_alltypes)

outputs_complete = list()
for (file in images){
	cat('processing', file)
  
  samplename = gsub('_cell_seg_data.txt', '', tail(strsplit(file, '/')[[1]],1))
  outputs_complete[[samplename]] = do_analyse(seg_path = file, PhenoOrder = phenotype_complete, 
                                          ColsOrder = color_complete, XposCol = 'Cell X Position', 
                                          YposCol = 'Cell Y Position', PhenoCol = 'Phenotype',
                      sample_name = samplename, plotter = plotter, fig.prefix = fig.prefix.complete,
                      r_vec = r_vec, spatstat_statistics = selected_spatstat_statistics)
}

```
The script `do_analyse` measured multiple statistics (`K` and `L` statistics were chosen in the example above) for the given image (`seg_path`).
Apart from the statistics itself, it will save images at the location specified by `fig.prefix`. 
One example image (K statistics) of all pairs of cell types considered (defined by `PhenoOrder`) is shown below. Note that `+` and `-` signs are converted to `T` and `F` for computation reasons. The red line and gray shadow indicate theoretical statistics when the pair of cell types have no geometric association. The solid black line represents the actual statistic.

![Example image](./complete/HO105-003_[11564,35679]/HO105-003_[11564,35679]_statistic_K.png)

```{r, eval=F}

phenotype_simple = list(Macrophage = c("CD163+PDL1-","CD163+PDL1+"),
                      Tcells = c("CD3+CD8-PD1-", "CD3+CD8-PD1+", "CD3+CD8+PD1-", "CD3+CD8+PD1+"),
                      Tumors = c("PAX5+PDL1-", "PAX5+PDL1+"),
                      Others = c("Other", "Other PDL1+")
                      )

color_simple = c(Macrophage = "magenta", Tcells = "red", Tumors = "orange", Others = "gray")

fig.prefix.simple = './simple'

outputs_simple = list()
for (file in images){
	cat('processing', file)
  
  samplename = gsub('_cell_seg_data.txt', '', tail(strsplit(file, '/')[[1]],1))
  outputs_simple[[samplename]] = do_analyse(seg_path = file, PhenoOrder = phenotype_simple, 
                                          ColsOrder = color_complete, XposCol = 'Cell X Position', 
                                          YposCol = 'Cell Y Position', PhenoCol = 'Phenotype',
                      sample_name = samplename, plotter = plotter, fig.prefix = fig.prefix.simple,
                      r_vec = r_vec, spatstat_statistics = selected_spatstat_statistics)
}


```


```{r HO105_extreme,fig.width=8, fig.height=8, eval=F} 
source('spatstat_vectra.R')
# source('C:/Users/erikb/Documents/Studie/Thesis - Local/spatstat_vectra_backup_2020-05-04/spatstat_vectra 2020-07-21.R')
extreme = file.path(base_path, c(
		'HO105-194_[16067,37856]_cell_seg_data.txt', # PD-L1 positive cluster + Tumor cluster
		'HO105-194_[15564,37856]_cell_seg_data.txt', # PD-L1 positive cluster + Tumor cluster
		'HO105-194_[16029,38205]_cell_seg_data.txt' # not clustered
	))
D = file.path(base_path, c(
    'HO105-006_[9554,40992]_cell_seg_data.txt',  # complete and simple: no component (not in marits file) 1
    'HO105-006_[10041,41957]_cell_seg_data.txt', # complete and simple: no component (not in marits file) 2
		'HO105-140_[16281,41985]_cell_seg_data.txt', # complete: Fest spacing  3 inside alltypes(..,fun=J) small counts 5,1
		'HO105-083_[20512,33844]_cell_seg_data.txt', # simple: (maybe counts) 4 inside alltypes no Tcells solved
		'HO105-083_[18166,32820]_cell_seg_data.txt', # simple: (maybe counts) 5 inside alltypes no Tcells solved
		'HO105-097_[9517,43652]_cell_seg_data.txt',  # simple: (maybe counts) 6 inside alltypes no Tcells solved
		'HO105-097_[8938,43999]_cell_seg_data.txt',  # simple: (maybe counts) 7 inside alltypes no Tcells solved
		'HO105-097_[11295,45384]_cell_seg_data.txt', # simple: (maybe counts) 8 inside alltypes no Tcells solved
		'HO105-121_[8016,47627]_cell_seg_data.txt',  # complete: Fest spacing 9 inside alltypes(..,fun=J) small counts 5,2
		'HO105-083_[20684,34346]_cell_seg_data.txt', # simple: (maybe counts) 10 inside alltypes no Tcells solved
		'HO105-083_[14332,34366]_cell_seg_data.txt', # simple: (maybe counts) 11 inside alltypes no Tcells solved
		'HO105-083_[14941,33993]_cell_seg_data.txt', # simple: (maybe counts) 12 inside alltypes no Tcells solved
		'HO105-083_[18456,34691]_cell_seg_data.txt', # simple: (maybe counts) 13 inside alltypes no Tcells solved
		'HO105-083_[15677,36542]_cell_seg_data.txt', # simple: (maybe counts) 14 inside alltypes no Tcells solved
		'HO105-083_[15726,34331]_cell_seg_data.txt', # simple: (maybe counts) 15 inside alltypes no Tcells solved
		'HO105-095_[15101,39877]_cell_seg_data.txt', # complete: Fest spacing 16 inside alltypes(..,fun=J) small counts 5,1 maybe area is to thin?
		'HO105-083_[16066,33773]_cell_seg_data.txt', # simple: (maybe counts) 17 inside alltypes no Tcells solved
		'HO105-083_[16367,35783]_cell_seg_data.txt'  # simple: (maybe counts) 18 inside alltypes no Tcells solved
	))
# extreme = D
# Plot variables

plotter_sample = T
plotter_quadrat = T
plotter_alltypes = T

plotter = list(plotter_sample, plotter_quadrat, plotter_alltypes)

# Select spatial statistics, or NULL then all inbuild spatial statistics are selected
selected_spatstat_statistics = list('K', "L") # 'ALL', NULL, list('Gdot', 'K') list("G","F", "J","Gdot", "Jdot", "K", "L", "pcf", "Kdot", "Ldot")

# debug

extreme = file.path(base_path,'HO105-083_[16577,32382]_cell_seg_data.txt') # Ma= 95, Ot= 129, Tc= 1, Tu= 732
extreme = file.path(base_path,'HO105-097_[9517,43652]_cell_seg_data.txt') # Ma= 37, Ot= 198, Tc= 0, Tu= 37
extreme = file.path(base_path,'HO105-006_[13675,45799]_cell_seg_data.txt') # Ma= 12, Ot= 386, Tc= 2, Tu= 720
extreme = file.path(base_path,'HO105-060_[19134,52452]_cell_seg_data.txt') # Ma= 21, Ot= 404, Tc= 3, Tu= 847
# extreme = file.path(base_path,'HO105-062_[11908,48533]_cell_seg_data.txt') # Ma= 7, Ot= 224, Tc= 4, Tu= 1360
# extreme = file.path(base_path,'HO105-101_[12806,43611]_cell_seg_data.txt') # Ma= 63, Ot= 405, Tc= 5, Tu= 1264
extreme = file.path(base_path,'HO105-083_[14332,34366]_cell_seg_data.txt') # 541 NA's  Ma =55 Ot=146 Tc=0 Tu=706

# names(features_raw[,'counts_sample_Tcells'][features_raw[,'counts_sample_Tcells'] == 0])
# View( Tc0 <-features_raw[names(features_raw[,'counts_sample_Tcells'][features_raw[,'counts_sample_Tcells'] == 0]),]); print(dim(Tc0));unique(colSums(is.na(Tc0))); 15
# names(features_raw[,'counts_sample_Tcells'][features_raw[,'counts_sample_Tumors'] == 0])

# extreme = c(extreme[3],D[8])

# extreme = file.path(base_path, paste0(c(threes_names[1]),'_cell_seg_data.txt'))

outputs_auto = list()
for (file in extreme){
	cat('processing', file)
  
  samplename = gsub('_cell_seg_data.txt', '', tail(strsplit(file, '/')[[1]],1))
	outputs_auto[[samplename]] = do_analyse(seg_path = file, PhenoOrder = phenotype_simple, ColsOrder = color_simple,
                      XposCol = 'Cell X Position', YposCol = 'Cell Y Position', PhenoCol = 'Phenotype',
                      sample_name = samplename, plotter = plotter, fig.prefix = fig.prefix.simple,
                      r_vec = r_vec, spatstat_statistics = selected_spatstat_statistics)
}

# saveRDS(outputs_auto, file = 'feature_matrix_output.RDS')
M = feature_extract(outputs = outputs_auto)
# write.csv(x = M,file = file.path(fig.prefix.simple, 'feature_matrix.csv'))
```


```{r marit HO105 image check  all via import, fig.width=8, fig.height=8, eval=F}
source('spatstat_vectra.R')
library('readxl')


#### Preprocessing ####
# Define path for all data
base_path <- '../Exported data tables/' # Working for Yongsoo
# base_path <- 'C:/Users/erikb/Documents/Thesis - Local/Data Marit voor Erik Bosch/Erik Bosch/Data complete phenotyping' # working for Erik
base_path = path.expand("~/Studie/Thesis - Local/Data Marit voor Erik Bosch/Erik Bosch/Data complete phenotyping")

# define path for selected images by Marit
image_check_path = NULL # working for Yongsoo (CHANGE)
image_check_path <- 'C:/Users/erikb/Documents/Studie/Thesis - Local/spatstat_vectra/HO105 image check.xlsx' # working for Erik
image_check_path <- path.expand('~/Studie/Thesis - Local/spatstat_vectra/HO105 tumor_border annotation.xlsx') # working for Erik


# all_files_x_ = rownames(base_path, pattern =  'HO105-',full.names = TRUE)
# names_with_x_ = list.files(base_path, pattern =  'HO105-[0-9]+_[0-9]_',full.names = TRUE)
# 
# if (length(names_with_x_)>0){
#     new_names = str_replace_all(all_files_x_, pattern = '_[0-9]_', replacement = '\\_')
#     warning(paste('file names have been renamed in the current directory',getwd()))
#     rename_log = file.rename(all_files_x_, new_names)
# }





#### Read file for selected images by Marit ####
HO105_image_check <- read_excel(image_check_path)
HOnrs_unfil = HO105_image_check$`HO105 nr`

tumor_images_data_unfil = HO105_image_check$`Tumor images`
# get the MSI's through regex finding: open bracket->MSI numbers->comma->MSI numbers-> closed bracket
tumor_images_unfil = str_extract_all(tumor_images_data_unfil,'\\[[0-9]+,[0-9]+\\]')
# create filter to check for non empty content
non_empty = !sapply(tumor_images_unfil,is_empty)
# create filtered tumor_images and HOnrs
tumor_images = tumor_images_unfil[non_empty]
HOnrs = HOnrs_unfil[non_empty]
names(tumor_images) = HOnrs
# create string with paths to segmentation data and component data to analyse
sample_names_marit = lapply(seq_along(HOnrs),function(x) paste0(names(tumor_images)[x],'_',unlist(tumor_images[HOnrs][x])))
sample_names_marit_str = unlist(sample_names_marit)
length(sample_names_marit_str)


border_images_data_unfil = HO105_image_check$`Border images`
# get the MSI's through regex finding: open closed bracket->MSI numbers->comma->MSI numbers-> closed closed bracket
border_images_unfil = str_extract_all(border_images_data_unfil,'\\[[0-9]+,[0-9]+\\]')
# create filter to check for non empty content
non_empty_border = !sapply(border_images_unfil,is_empty)
# create filtered tumor_images and HOnrs
border_images = border_images_unfil[non_empty_border]
HOnrs_border = HOnrs_unfil[non_empty_border]
names(border_images) = HOnrs_border

# #### check if the component data and segmentation data are one to one ####
# # # find all files of HO105
# # all_files = list.files(base_path, pattern =  'HO105-')
# 
# ## find all available component data of HO105
# available_comp_tif = list.files(base_path, pattern = '_component_data.tif')
# # find the sample names of all available component data
# available_comp = lapply(available_comp_tif, function(x) gsub('_component_data.tif','',x))
# available_comp_str = unlist(available_comp)
# length(available_comp_str)
# 
# ## find all available segmentation data of HO105
# available_seg_txt = list.files(base_path, pattern = '_cell_seg_data.txt')
# # find the sample names of the all available segmentation data
# available_seg = lapply(available_seg_txt, function(x) gsub('_cell_seg_data.txt','',x))
# available_seg_str = unlist(available_seg)
# length(available_seg_str)


# # take intersection to retrieve the matching names with the data from Marits file
# matching_names_marit = intersect(sample_names_marit_str,available_comp_str)
# length(matching_names_marit)
# 
# # take the difference to retrieve the sample names of the missing component data
# diff_available_comp = setdiff(sample_names_marit_str,available_comp_str)
# length(diff_available_comp)
# 
# #take set difference of names of actual available components and names of components of 
# diff_available_seg = setdiff(sample_names_marit_str,available_seg_str)
# length(diff_available_seg)




```

```{r clinical vs prediction data 22nd Jun, eval=F}
source('spatstat_vectra.R')

# import simplified matrix

feature_matrix_simple_22 = readRDS(path.expand('~/Studie/Thesis - Local/Extracted_Features_22ndJun.RDS'))

cols_without_J = str_subset(colnames(feature_matrix_simple_22),'J_radius', negate = TRUE)
cols_without_Jdot = str_subset(colnames(feature_matrix_simple_22),'Jdot_radius', negate = TRUE)
cols_without_J_Total = intersect(cols_without_J, cols_without_Jdot)
feature_matrix_simple_22_without_J = feature_matrix_simple_22[,cols_without_J_Total]

# View(rowSums((is.na(feature_matrix_simple_22_without_J))))
# View(colSums((is.na(feature_matrix_simple_22_without_J)))) # 22 = 14 + 8 times MED_min_Tcells_Tcells == NA
# View(colSums((feature_matrix_simple_22_without_J ==0))) # 14 times 0 Tcells
# View(colSums((feature_matrix_simple_22_without_J ==1))) # 8 times 1 Tcell

NAs_per_sample = rowSums((is.na(feature_matrix_simple_22_without_J)))
# View(feature_matrix_simple_22_without_J[rowSums((is.na(feature_matrix_simple_22_without_J)))>0,])
df_and_NAs = cbind(NAs_per_sample,feature_matrix_simple_22_without_J)
View(df_and_NAs)

# 541 NAs is when a phenotype count has 0
# 26 NAs is when a phenotype count has 1
# 6 and 3 NAs is when a phenotype as extensively many counts.

# samples with 3 NA's have loooooooots of TUmors and results in a Fstatistic of NA for large radi: 30,40,50
threes = df_and_NAs[threes_names <- c('HO105-033_[20154,47207]','HO105-183_[10192,36437]', 'HO105-183_[11016,38792]', 'HO105-183_[15933,32458]', 'HO105-183_[16366,40912]', 'HO105-183_[18063,33897]',  'HO105-183_[9034,45388]'),] # test for F30,40,50 ook normalized geeft daardoor 0
# samples with 6 NA's have
sixes = df_and_NAs[sixes_names <- 'HO105-061_[11807,41972]',] # test for F30,40,50 ook normalized geeft daardoor 0

# samples with 26 NA's have a count of 1 for a phenotype
stop()
#### renaming rownames in simplified matrix with '_1_', '_2_', '_3_' and '_4_' (replace with '_') ####
all_rownames = rownames(feature_matrix_simple_22)
all_rownames = str_replace_all(all_rownames, pattern = '\\_[0-9]\\_', replacement = '\\_')

rownames(feature_matrix_simple_22) = all_rownames


# match_indices = match(sample_names_marit_str,all_rownames)
# feature_matrix_simple_22_tumor = feature_matrix_simple_22[match_indices,]
# 
# saveRDS(feature_matrix_simple_22_tumor, file = 'feature_matrix_simple_22_tumor.RDS')
# stop('joe')



mat_simple_allpatients_complete_cases_mean = statisticPerPatient(feature_matrix_simple_22, statistic = 'mean', na.handler = 'complete_cases')
# apply(feature_matrix_simple_22,1,function(x) sum(is.na(x))) if all bigger than 0, then every sample has NA's in them, so no complete cases exist
mat_simple_allpatients_ignore_na_mean = statisticPerPatient(feature_matrix_simple_22, statistic = 'mean', na.handler = 'ignore_na')
# apply(mat_simple_allpatients_ignore_na_mean,1,function(x) sum(is.na(x))) NA's are not a problem in this setting

mat_simple_allpatients_complete_cases_median = statisticPerPatient(feature_matrix_simple_22, statistic = 'median', na.handler = 'complete_cases')
# apply(feature_matrix_simple_22,1,function(x) sum(is.na(x))) if all bigger than 0, then every sample has NA's in them, so no complete cases exist
mat_simple_allpatients_ignore_na_median = statisticPerPatient(feature_matrix_simple_22, statistic = 'median', na.handler = 'ignore_na') 



#### finding the the widths for significance band of each feature per patient

# feature_matrix_simple_22 = M
uni_stats = unique(str_subset(colnames(feature_matrix_simple_22), '_radius'))


spats_unnorm = str_subset(uni_stats, 'Normalized',negate = TRUE)
m = matrix(NA,nrow = length(rownames(feature_matrix_simple_22)), ncol = length(spats_unnorm))
rownames(m) = rownames(feature_matrix_simple_22)
colnames(m) = spats_unnorm

for (patient in rownames(feature_matrix_simple_22)){
  for (stat in spats_unnorm){
    unnorm = feature_matrix_simple_22[patient,stat]
    nnorm = feature_matrix_simple_22[patient,paste0('Normalized_',stat)]
    width_sign_band = unnorm/nnorm
    m[patient,stat] = width_sign_band
  }
}

width_sign_band_min = apply(m,1,function(x) min(x,na.rm = TRUE))
width_sign_band_max = apply(m,1,function(x) max(x,na.rm = TRUE))

# min(width_sign_band_min) = 2.556935e-05
# max(width_sign_band_max) = 172298.4


# m = matrix(1:36,nrow = 6)
# rownames(m) = list('HO105-097_[9517,43652]', 'HO105-097_[8938,43999]', 'HO105-097_[11295,45384]', 'HO105-121_[8016,47627]', 'HO105-083_[20684,34346]', 'HO105-083_[14332,34366]')
# colnames(m) = letters[1:6]
# m[4] = NA
# mat = m

# NaN -> apply(matrix(c(NA,1:4),nrow = 1),2, mean, na.rm = TRUE)


```

```{r clinical vs prediction data 26nd Jun, eval=F}
source('spatstat_vectra.R')

# import simplified matrix

feature_matrix_simple_26 = readRDS(path.expand('~/Studie/Thesis - Local/Extracted_Features_26ndJun.RDS'))

NAs_per_sample_all_cols = rowSums((is.na(feature_matrix_simple_26)))
table(NAs_per_sample_all_cols)
# NAs_per_sample_all_cols
#   0   3   4   7   8  10  11  12  15  16  20  24  26  30  34  38 669 672 675 
# 443   1 334   5 220   1   1 103   1  31  16   7   4   2   1   2   6   5   3
df_and_NAs_all_colls = cbind(NAs_per_sample_all_cols,feature_matrix_simple_26)
View(df_and_NAs_all_colls)


cols_without_J = str_subset(colnames(feature_matrix_simple_26),'J_radius', negate = TRUE)
cols_without_Jdot = str_subset(colnames(feature_matrix_simple_26),'Jdot_radius', negate = TRUE)
cols_without_J_Total = intersect(cols_without_J, cols_without_Jdot)
feature_matrix_simple_26_without_J = feature_matrix_simple_26[,cols_without_J_Total]


NAs_per_sample = rowSums((is.na(feature_matrix_simple_26_without_J)))
table(NAs_per_sample)
#    0    3    6   26  541 
# 1155    7    1    9   14
# View(feature_matrix_simple_26_without_J[rowSums((is.na(feature_matrix_simple_26_without_J)))>0,])
df_and_NAs = cbind(NAs_per_sample,feature_matrix_simple_26_without_J)
View(df_and_NAs)

# 541 NAs is when a phenotype count has 0
# 26 NAs is when a phenotype count has 1
# 6 and 3 NAs is when a phenotype has extensively many counts (over 4000)

# samples with 3 NA's have loooooooots of TUmors and results in a Fstatistic of NA for large radi: 30,40,50
threes = df_and_NAs[threes_names <- c('HO105-033_[20154,47207]','HO105-183_[10192,36437]', 'HO105-183_[11016,38792]', 'HO105-183_[15933,32458]', 'HO105-183_[16366,40912]', 'HO105-183_[18063,33897]',  'HO105-183_[9034,45388]'),str_subset(colnames(feature_matrix_simple_26),'F')] # test for F30,40,50 ook normalized geeft daardoor 0
# samples with 6 NA's have
sixes = df_and_NAs[sixes_names <- 'HO105-061_[11807,41972]',str_subset(colnames(feature_matrix_simple_26),'F')] # test for F30,40,50 ook normalized geeft daardoor 0

# samples with 26 NA's have a count of 1 for a phenotype

# sum(feature_matrix_simple_22 == feature_matrix_simple_26,na.rm = TRUE)+sum(is.na(feature_matrix_simple_22)) = 1186*1686 so both matrices are the same.

```

```{r clinical vs prediction data 30nd Jun, eval=F}
source('spatstat_vectra.R')

# import simplified matrix

feature_matrix_simple_30 = readRDS(path.expand('~/Studie/Thesis - Local/Extracted_Features_30thJun.RDS'))

NAs_per_sample_all_cols = rowSums((is.na(feature_matrix_simple_30)))
# table(NAs_per_sample_all_cols)
# NAs_per_sample_all_cols
#   0   4   8  12  16  20  24  25  26  28  29  30  32  36  37  38  40  41  42  43  44  48  52  54  56  58  60  62  64  66 669 678 684 687 690 693 
# 248  56  23  60 154 106  98   1   1 108   1   1 152  77   1   3  48   2   1   1  14   4   2   1   2   1   2   2   1   1   4   3   1   2   3   1 
# Quite a few NA's because of J-statistic
df_and_NAs_all_colls = cbind(NAs_per_sample_all_cols,feature_matrix_simple_30)



cols_without_J = str_subset(colnames(feature_matrix_simple_30),'J_radius', negate = TRUE)
cols_without_Jdot = str_subset(colnames(feature_matrix_simple_30),'Jdot_radius', negate = TRUE)
cols_without_J_Total = intersect(cols_without_J, cols_without_Jdot)
feature_matrix_simple_30_without_J = feature_matrix_simple_30[,cols_without_J_Total]


NAs_per_sample = rowSums((is.na(feature_matrix_simple_30_without_J)))
table(NAs_per_sample)
#    0   26  541 
# 1163    9   14
# Without J-statistic it is more consistent

# 541 NAs is when a phenotype count has 0, expected outcome
# 26 NAs is when a phenotype count has 1, expected outcome

View(feature_matrix_simple_30_without_J)

saveRDS(feature_matrix_simple_30_without_J,file = path.expand('~/Studie/Thesis - Local/Extracted_Features_1stJuly_Simple.RDS'))

```

